---
title: "Practicas general"
author: "Ignacio Borlaf Mena"
date: "17/1/2022"
output: html_document
---

```{r setup, include=FALSE}
# install.packages("palmerpenguins")
knitr::opts_chunk$set(echo = TRUE)
library(palmerpenguins)
library(dplyr)
library(tidyr)
library(ggplot2)
data(penguins, package = 'palmerpenguins')
```

## Number of individuals

### Per species

```{r spec_counts, warn=FALSE}
penguins |>
  group_by(species) |>
  summarize(count=n())
```
### Per species and sex

```{r spec_counts_sex, warn=FALSE}
penguins |>
  filter(!is.na(sex)) |>
  group_by(species, sex) |>
  summarize(count=n())
```

### Summary statistics

#### DPLYR, output single-row tibble

```{r stats_dplyr_across}

col_vector <- c("bill_length_mm", "bill_depth_mm",
                "flipper_length_mm", "body_mass_g")
sep <- "@"
naming <- paste("{.col}", "{.fn}", sep = sep)
stat_list <- list(min=min, mean=mean, sd=sd, max=max)

penguins |>
  summarise(across(col_vector, stat_list, na.rm=TRUE, .names = naming)) |>
  pivot_longer(everything(), names_to = c("var", ".value"), names_sep = sep)
```

## Visualiztion

### Barplot with the by-species count of individuals by island

```{r barplot}
color_list <- list(Chinstrap="purple", Gentoo="blue", Adelie="orange")
ggplot(data = penguins) +
  geom_bar(aes(x=island, fill=species)) +
  scale_fill_manual(values = color_list)
  
```

### Which species have been found in all three islands?

Adelie

### How many individuals have been sampled in the Dream island?

```{r spec_counts_island}
penguins |>
  group_by(island) |>
  count() |>
  filter(island == "Dream") |>
  select(n)
```
124

### Boxplot
```{r boxplot}
penguins |>
  filter(!is.na(sex)) |>
  ggplot() +
    geom_boxplot(aes(x=sex, y=body_mass_g)) +
    facet_grid(cols = vars(species)) +
    labs(x = "Sex", y = "Body mass (g)")
```

* Gentoo has the largest body weight
* Chinstrap has the smalles difference in body weight between sexes